---
title: "Base Case Analysis"
---

# Base Case Analysis

This analysis examines the distribution of outputs from multiple simulation runs at the same parameter values (base case).

## Load Data

```{r}
# Load the parameter map and model outputs
param_map <- read.table("../data/sim_modeloutputs.2025.Dec.12.11_22_51.batch_param_map.txt",
                        header = TRUE,
                        sep = ",",
                        check.names = FALSE,
                        stringsAsFactors = TRUE)

model_outputs <- read.table("../data/sim_modeloutputs.2025.Dec.12.11_22_51.txt",
                            header = TRUE,
                            sep = ",",
                            stringsAsFactors = TRUE)

# Join the datasets on "run"
combined_data <- merge(param_map, model_outputs, by = "run")

# Filter out runs 0-5
combined_data <- combined_data[combined_data$run > 5, ]

# Display basic information about the dataset
cat(sprintf("Number of runs: %d (after filtering out runs 0-5)\n", nrow(combined_data)))
cat(sprintf("Number of variables: %d\n\n", ncol(combined_data)))

# Show column names and types
cat("Variables in the dataset:\n")
str(combined_data)
```

## Number of Transmissions by Run Number

```{r}
# Scatter plot of NumberOfTransmissions vs run number
plot(combined_data$run, combined_data$NumberOfTransmissions,
     xlab = "Run Number",
     ylab = "Number of Transmissions",
     main = "Number of Transmissions by Run Number",
     pch = 19,
     col = "steelblue")

# Add horizontal line at the mean
abline(h = mean(combined_data$NumberOfTransmissions), col = "red", lwd = 2, lty = 2)
legend("topright",
       legend = sprintf("Mean: %.1f", mean(combined_data$NumberOfTransmissions)),
       col = "red", lty = 2, lwd = 2)
```

## Investigating Outliers (Parallel Computing Issue)

```{r}
# Find outliers in NumberOfTransmissions
cat("Investigating outlier runs:\n")
cat("===========================\n\n")

# Calculate z-scores
mean_trans <- mean(combined_data$NumberOfTransmissions)
sd_trans <- sd(combined_data$NumberOfTransmissions)
combined_data$z_score <- (combined_data$NumberOfTransmissions - mean_trans) / sd_trans

# Sort by run number to check if first 10 are outliers
sorted_by_run <- combined_data[order(combined_data$run), ]

# Check statistics for first 10 vs rest
first_10 <- sorted_by_run[1:min(10, nrow(sorted_by_run)), ]
rest <- sorted_by_run[(min(11, nrow(sorted_by_run))+1):nrow(sorted_by_run), ]

cat("Comparing first 10 runs vs. remaining runs:\n")
cat("============================================\n\n")

cat("First 10 runs (potentially distributed to cores at startup):\n")
cat(sprintf("  Mean Transmissions: %.1f (SD: %.1f)\n",
            mean(first_10$NumberOfTransmissions),
            sd(first_10$NumberOfTransmissions)))
cat(sprintf("  Range: %d to %d\n",
            min(first_10$NumberOfTransmissions),
            max(first_10$NumberOfTransmissions)))

if(nrow(rest) > 0) {
  cat("\nRemaining runs:\n")
  cat(sprintf("  Mean Transmissions: %.1f (SD: %.1f)\n",
              mean(rest$NumberOfTransmissions),
              sd(rest$NumberOfTransmissions)))
  cat(sprintf("  Range: %d to %d\n",
              min(rest$NumberOfTransmissions),
              max(rest$NumberOfTransmissions)))

  # T-test to check if means are significantly different
  t_result <- t.test(first_10$NumberOfTransmissions, rest$NumberOfTransmissions)
  cat(sprintf("\nT-test p-value: %.6f\n", t_result$p.value))
  if(t_result$p.value < 0.05) {
    cat("** SIGNIFICANT DIFFERENCE detected (p < 0.05) **\n")
    cat("This suggests a parallel computing initialization issue!\n")
  }
}

cat("\n\nRun-by-run breakdown of first 15 runs:\n")
cat("Run | ExtraIter | Transmissions | Z-score | Inferred Core\n")
cat("----|-----------|---------------|---------|---------------\n")
for(i in 1:min(15, nrow(sorted_by_run))) {
  inferred_core <- (sorted_by_run$run[i] - 1) %% 10
  cat(sprintf("%3d | %9d | %13d | %7.2f | %13d\n",
              sorted_by_run$run[i],
              sorted_by_run$extraIteration[i],
              sorted_by_run$NumberOfTransmissions[i],
              sorted_by_run$z_score[i],
              inferred_core))
}

# Check if there's a pattern by inferred core
cat("\n\nAnalysis by inferred core (assuming run %% 10 mapping):\n")
cat("=========================================================\n")
sorted_by_run$inferred_core <- (sorted_by_run$run - 1) %% 10

core_stats <- aggregate(NumberOfTransmissions ~ inferred_core,
                       data = sorted_by_run,
                       FUN = function(x) c(mean = mean(x),
                                          sd = sd(x),
                                          n = length(x)))

cat("\nCore | Mean Trans | SD      | N runs\n")
cat("-----|------------|---------|-------\n")
for(i in 0:9) {
  if(i %in% core_stats$inferred_core) {
    idx <- which(core_stats$inferred_core == i)
    cat(sprintf("%4d | %10.1f | %7.1f | %6d\n",
                i,
                core_stats$NumberOfTransmissions[idx,1],
                core_stats$NumberOfTransmissions[idx,2],
                core_stats$NumberOfTransmissions[idx,3]))
  }
}

# Check relationship between extraIteration and run number
cat("\n\nRelationship between run number and extraIteration:\n")
cat("====================================================\n")
cat("Run | ExtraIter | Run %% 10 | ExtraIter %% 10\n")
cat("----|-----------|----------|----------------\n")
for(i in 1:min(20, nrow(sorted_by_run))) {
  cat(sprintf("%3d | %9d | %8d | %15d\n",
              sorted_by_run$run[i],
              sorted_by_run$extraIteration[i],
              sorted_by_run$run[i] %% 10,
              sorted_by_run$extraIteration[i] %% 10))
}
```

## Importation Rate vs Importation Prevalence

```{r}
# Scatter plot of importationRate vs ImportationPrevalence
plot(combined_data$importationRate, combined_data$ImportationPrevalence,
     xlab = "Importation Rate (Parameter)",
     ylab = "Importation Prevalence (Output)",
     main = "Importation Rate vs Importation Prevalence",
     pch = 19,
     col = "darkblue")

# Add correlation information
cor_val <- cor(combined_data$importationRate, combined_data$ImportationPrevalence)
text(x = max(combined_data$importationRate) * 0.95,
     y = min(combined_data$ImportationPrevalence) * 1.05,
     labels = sprintf("Correlation: %.4f", cor_val),
     adj = c(1, 0))
```

## Output Distributions

### Clinical Detections

```{r}
hist(combined_data$ClinicalDetections,
     main = "Distribution of Clinical Detections",
     xlab = "Clinical Detections",
     col = "steelblue",
     breaks = 20)
abline(v = mean(combined_data$ClinicalDetections), col = "red", lwd = 2, lty = 2)
abline(v = median(combined_data$ClinicalDetections), col = "darkgreen", lwd = 2, lty = 2)
legend("topright",
       legend = c(sprintf("Mean: %.1f", mean(combined_data$ClinicalDetections)),
                  sprintf("Median: %.1f", median(combined_data$ClinicalDetections))),
       col = c("red", "darkgreen"), lty = 2, lwd = 2)
```

### Mean Daily Prevalence

```{r}
hist(combined_data$MeanDailyPrevalence,
     main = "Distribution of Mean Daily Prevalence",
     xlab = "Mean Daily Prevalence",
     col = "coral",
     breaks = 20)
abline(v = mean(combined_data$MeanDailyPrevalence), col = "red", lwd = 2, lty = 2)
abline(v = median(combined_data$MeanDailyPrevalence), col = "darkgreen", lwd = 2, lty = 2)
legend("topright",
       legend = c(sprintf("Mean: %.4f", mean(combined_data$MeanDailyPrevalence)),
                  sprintf("Median: %.4f", median(combined_data$MeanDailyPrevalence))),
       col = c("red", "darkgreen"), lty = 2, lwd = 2)
```

### Number of Transmissions

```{r}
hist(combined_data$NumberOfTransmissions,
     main = "Distribution of Number of Transmissions",
     xlab = "Number of Transmissions",
     col = "lightgreen",
     breaks = 20)
abline(v = mean(combined_data$NumberOfTransmissions), col = "red", lwd = 2, lty = 2)
abline(v = median(combined_data$NumberOfTransmissions), col = "darkgreen", lwd = 2, lty = 2)
legend("topright",
       legend = c(sprintf("Mean: %.1f", mean(combined_data$NumberOfTransmissions)),
                  sprintf("Median: %.1f", median(combined_data$NumberOfTransmissions))),
       col = c("red", "darkgreen"), lty = 2, lwd = 2)
```

### Mean Discharge Prevalence

```{r}
hist(combined_data$MeanDischargePrevalence,
     main = "Distribution of Mean Discharge Prevalence",
     xlab = "Mean Discharge Prevalence",
     col = "plum",
     breaks = 20)
abline(v = mean(combined_data$MeanDischargePrevalence), col = "red", lwd = 2, lty = 2)
abline(v = median(combined_data$MeanDischargePrevalence), col = "darkgreen", lwd = 2, lty = 2)
legend("topright",
       legend = c(sprintf("Mean: %.4f", mean(combined_data$MeanDischargePrevalence)),
                  sprintf("Median: %.4f", median(combined_data$MeanDischargePrevalence))),
       col = c("red", "darkgreen"), lty = 2, lwd = 2)
```

### Importation Prevalence

```{r}
hist(combined_data$ImportationPrevalence,
     main = "Distribution of Importation Prevalence",
     xlab = "Importation Prevalence",
     col = "khaki",
     breaks = 20)
abline(v = mean(combined_data$ImportationPrevalence), col = "red", lwd = 2, lty = 2)
abline(v = median(combined_data$ImportationPrevalence), col = "darkgreen", lwd = 2, lty = 2)
legend("topright",
       legend = c(sprintf("Mean: %.4f", mean(combined_data$ImportationPrevalence)),
                  sprintf("Median: %.4f", median(combined_data$ImportationPrevalence))),
       col = c("red", "darkgreen"), lty = 2, lwd = 2)
```

## Summary Statistics

```{r}
# Create a summary table for all numeric output variables
output_vars <- c("ClinicalDetections", "MeanDailyPrevalence", "MeanDischargePrevalence",
                 "ImportationPrevalence", "NumberOfTransmissions")

summary_table <- data.frame(
  Variable = output_vars,
  Mean = sapply(output_vars, function(v) mean(combined_data[[v]])),
  Median = sapply(output_vars, function(v) median(combined_data[[v]])),
  SD = sapply(output_vars, function(v) sd(combined_data[[v]])),
  Min = sapply(output_vars, function(v) min(combined_data[[v]])),
  Max = sapply(output_vars, function(v) max(combined_data[[v]])),
  CV = sapply(output_vars, function(v) sd(combined_data[[v]]) / mean(combined_data[[v]]))
)

# Format the table
print(summary_table, row.names = FALSE, digits = 4)
```
